Type Flattening Obfuscation

### 2020-01-29 paper/tool.tex
update

\section{Type flattening compiler}
We present \emph{uCc}, an open source C compiler that explicitly obfuscates high-level types, its
objective is to make reconstructing types from machine codes harder.
It is implemented in Rust, the source code and a brief user guide are given at (cite).
Our effort is to attack the core of type reconstruction algorithms,
these are the data flow used to build type constraints, and the \emph{semantic gap} between types in the high-level language and their
representations in the low-level machine code.
Beside custom tricks, \emph{uCc} uses extensively
obfuscation transformations based on Mixed Boolean Arithmetic expressions~\cite{eyrolles_obfuscation_2017,zhou_information_2007}.
To make some code diversity, the compiler is \emph{probabilistic}: given a source
code, \emph{uCc} generates each time a different (but computationally equivalent) output machine code.

@@

\begin{table}
	\begin{center}
		\caption{Functions and original types}\label{description}
		\begin{tabular}{|l|l|}
			\hline
			Description & Function type \\
			\hline
			identity & \texttt{int id(int)} \\
			\hline
			division & \texttt{int div(short, char)} \\
			\hline
			modular & \texttt{char mod(short, char)} \\
			\hline
			increase $p$ then dereference & \texttt{int inc\_deref(int *p)} \\
			\hline
			strlen  & \texttt{int slen(char*)} \\
			\hline
			sum of array $a$ & \texttt{long sum(short *a, int n)} \\
			\hline
			$n$-th fibonaci number & \texttt{int fibo(int n)} \\
			\hline
			num. of steps until $n \rightsquigarrow 1$ (Collatz's conj.)  & \texttt{int collatz(int n)} \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

\begin{table}
	\begin{adjustwidth}{-6.21in}{-5.95in}
		\begin{center}
			% \centering
			\caption{Type flattening effect}\label{flattening}
			% \begin{adjustbox}{max width=1.5\textwidth}
			\begin{tabular}{|l|l|l|l|l|}
				\hline
				Original type & Hex-Rays & JEB & Ghidra & Snowman \\
				\hline
				\texttt{int id(int)} & \texttt{int64 id(int64)} & \texttt{ulong id(ulong)} &
				\texttt{ulong id(void)} & \texttt{int64 id(int64)} \\
				\hline
				\texttt{short div(short, char)} & \texttt{int64 div(uint64, uint64)} & \texttt{div\_t div(int, int)} &
				\texttt{div\_t div(int, int)} & \texttt{int64\_t div(int64, int64)} \\
				\hline
				\texttt{char mod(short, char)} & \texttt{int64 mod(int64, int64)} & \texttt{ulong mod(ulong, ulong)} &
				\texttt{ulong mod(void)} & \texttt{int64 mod(int64, int64)} \\
				\hline
				\texttt{int inc\_deref(int*)} & \texttt{int64 inc\_deref(int64)} & \texttt{ulong inc\_deref(ulong)} &
				\texttt{ulong inc\_deref(undefined8)} & \texttt{int64 inc\_deref(uint64)} \\
				\hline
				\texttt{int slen(char*)} & \texttt{int64 slen(int64)} & \texttt{ulong slen(ulong)} &
				\texttt{ulong slen(undefined8)} & \texttt{int64 slen(uint64)} \\
				\hline
				\texttt{long sum(short*, int)} & \texttt{int64 sum(int64, uint64)} & \texttt{ulong sum(ulong, ulong)} & \texttt{undefined8 sum(undefined8)} & sn \\
				\hline
				\texttt{int fibo(int)} & \texttt{int64 fibo(uint64)} & \texttt{ulong fibo(ulong)} &
				\texttt{ulong fibo(undefined8)} & \texttt{int64 fibo(int64)} \\
				\hline
				\texttt{int collatz(int)} & \texttt{int64 collatz(int64)} & \texttt{ulong collatz(ulong)} &
				\texttt{ulong collatz(undefined8)} & \texttt{int64 collatz(int64)} \\
				\hline
			\end{tabular}
		\end{center}
	\end{adjustwidth}
\end{table}

>>>>

\section{Type flattening compiler}
We present \emph{uCc}, an open source C compiler that explicitly obfuscates high-level types, \emph{uCc}'s
objective is to make reconstructing C types from machine codes harder.
The compiler is implemented in Rust, the source code and a brief user guide are given at (cite); different
from other code obfuscation implementations, \emph{uCc} uses cranelift~\cite{noauthor_cranelift_nodate} as
backend.

Our effort is to attack the core of type reconstruction algorithms,
these are the data flow used to build type constraints, and the \emph{semantic gap} between types in the high-level language and their
representations in the low-level machine code.
Beside custom tricks which would be fully presented in another paper, \emph{uCc} uses extensively
obfuscation transformations based on Mixed Boolean Arithmetic expressions~\cite{eyrolles_obfuscation_2017,zhou_information_2007}.
To make some code diversity, the compiler is \emph{probabilistic}: given a source
code, \emph{uCc} generates each time a different (but computationally equivalent) output machine code.

@@

\begin{table}
	\begin{center}
		\caption{Functions and original types}\label{description}
		\begin{tabular}{|l|l|}
			\hline
			Description & Function type \\
			\hline
			identity & \texttt{int id(int)} \\
			\hline
			division & \texttt{int div(short, char)} \\
			\hline
			modular & \texttt{char mod(short, char)} \\
			\hline
			increase $p \leftarrow p + 1$ then dereference & \texttt{int inc\_deref(int *p)} \\
			\hline
			strlen  & \texttt{int slen(char* s)} \\
			\hline
			sum of array $a$ & \texttt{long sum(short *a, int n)} \\
			\hline
			$n$-th fibonaci number (recursive impl.) & \texttt{int fibo(int n)} \\
			\hline
			num. of steps until $n \rightsquigarrow 1$ (Collatz's conj.)  & \texttt{int collatz(int n)} \\
			\hline
			sdbm hash & \texttt{int sdbm(char *s)} \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

\begin{table}
	\begin{adjustwidth}{-6.21in}{-5.95in}
		\begin{center}
			% \centering
			\caption{Type flattening effect}\label{flattening}
			% \begin{adjustbox}{max width=1.5\textwidth}
			\begin{tabular}{|l|l|l|l|l|}
				\hline
				Original type & Hex-Rays & JEB & Ghidra & Snowman \\
				\hline
				\texttt{int id(int)} & \texttt{int64 id(int64)} & \texttt{ulong id(ulong)} &
				\texttt{ulong id(void)} & \texttt{int64 id(int64)} \\
				\hline
				\texttt{short div(short, char)} & \texttt{int64 div(uint64, uint64)} & \texttt{div\_t div(int, int)} &
				\texttt{div\_t div(int, int)} & \texttt{int64\_t div(int64, int64)} \\
				\hline
				\texttt{char mod(short, char)} & \texttt{int64 mod(int64, int64)} & \texttt{ulong mod(ulong, ulong)} &
				\texttt{ulong mod(void)} & \texttt{int64 mod(int64, int64)} \\
				\hline
				\texttt{int inc\_deref(int*)} & \texttt{int64 inc\_deref(int64)} & \texttt{ulong inc\_deref(ulong)} &
				\texttt{ulong inc\_deref(undefined8)} & \texttt{int64 inc\_deref(uint64)} \\
				\hline
				\texttt{int slen(char*)} & \texttt{int64 slen(int64)} & \texttt{ulong slen(ulong)} &
				\texttt{ulong slen(undefined8)} & \texttt{int64 slen(uint64)} \\
				\hline
				\texttt{long sum(short*, int)} & \texttt{int64 sum(int64, uint64)} & \texttt{ulong sum(ulong, ulong)} & \texttt{undefined8 sum(undefined8)} & sn \\
				\hline
				\texttt{int fibo(int)} & \texttt{int64 fibo(uint64)} & \texttt{ulong fibo(ulong)} &
				\texttt{ulong fibo(undefined8)} & \texttt{int64 fibo(int64)} \\
				\hline
				\texttt{int collatz(int)} & \texttt{int64 collatz(int64)} & \texttt{ulong collatz(ulong)} &
				\texttt{ulong collatz(undefined8)} & \texttt{int64 collatz(int64)} \\
				\hline
				\texttt{int sdbm(char*)} & \texttt{int64 sdbm(uint64)} & jeb & \texttt{ulong sdbm(undefined8)} & sn \\
				\hline
			\end{tabular}
		\end{center}
	\end{adjustwidth}
\end{table}
